Binning Strategies and Related Loss for Binned Scatterplots of Large Data

### 2016-12-03 versionForJCGS/PaperJCGS.Rnw
Small follow up wording edit to Scott MISE description.

Note that while the losses from creating a binned scatterplot may turn out to be substantial, they present a huge gain with respect to an traditional scatterplot, where density information is implicitly masked in large data situations. The idea of loss from one-dimensional binning was explored by Scott using mean integrated squared error as the loss function to be optimized by the choice of the number of bins in  the construction of histograms \citep{scott1979}. He later extended this discussion to two-dimensional binning, where he compared the mean integrated squared error (MISE) loss for hexagonal, rectangular and triangular binning; finding that hexagonal and rectangular binning performed similarly, both far superior to triangular binning \citep{scott1992}. \ktm{Scott's MISE assesses the difference between the empirical density and the binned density approximation, as integrated the squared deviation over $\mathbb{R}^2$. Whereas, we quantify binning losses using the spatial displacement of the points to bin centers using euclidean distances, which better reflects the visual information lost in the shift from the traditional to binned scatterplots.}

>>>>

Note that while the losses from creating a binned scatterplot may turn out to be substantial, they present a huge gain with respect to an traditional scatterplot, where density information is implicitly masked in large data situations. The idea of loss from one-dimensional binning was explored by Scott using mean integrated squared error as the loss function to be optimized by the choice of the number of bins in  the construction of histograms \citep{scott1979}. He later extended this discussion to two-dimensional binning, where he compared the mean integrated squared error (MISE) loss for hexagonal, rectangular and triangular binning; finding that hexagonal and rectangular binning performed similarly, both far superior to triangular binning \citep{scott1992}. \ktm{Scott's MISE assesses the difference between the empirical density and the binned density approximation, via integration of the squared deviation over $\mathbb{R}^2$. Whereas, we quantify binning losses using the spatial displacement of the points to bin centers using euclidean distances, which better reflects the visual information lost in the shift from the traditional to binned scatterplots.}
